Robot Team III: Scotty Strikes Back

### 2024-04-15 paper/paper.tex
updated text for intro

1. Classification is a common remote sensing and hyperspectral imaging problem
    - Endmember Identification and Spectral Unmixing is sub-problem specific to hyperspectral data

2. HSI Size allows their incorporation onto drones
    - \cite{robot-team-1}
    - \cite{robot-team-2}

3. Provided with in situ measurements, HSI can be used to predict a variety of relevant water quality and water composition parameters. Here we can describe our previous work

4. However, we would like to be able to identify the presence of spectral signatures for which we do not already have an in situ sensor.

5. Unsupervised classification techniques are nice however standard algorithms like k-means clustering require prior knowledge of the number of constituent endmembers and do not, in general, provide any interpretable relationship between classes e.g. how is class 1 more similar to class 2 than to class 10. The Self Organizing map introduced by Kohonen is a popular technique which enables an unsupervised classification of data where classes

- We should make a note about Autoencoders and VAEs here. While they work well in many contexts, they do not provide a relationship between classes and there is no obvious way to choose the size of the encoding

6. "In this paper, we extend our previous work into the unsupervised classification domain..." Here we demonstrate the application of Generative Topographic mapping which addresses these issues by providing a probability distribution of resulting classes.

7. We apply the GTM to datasets collected by our Robot Team specifically showcasing the identification of spectral signatures for Rhodamine Dye, Algae, etc...

8. With the signatures identified, class abundance, and therefore area estimation, can be performed rapidly, in real time, by utilizing the learned signatures with the nromalized spectral similarity score (or other similar metrics).

IDEA: spectral unmixing of water is challenging due to nonlinear effects of scattering by solids and absorption by dissolved chemicals (dyes etc...) this motivates a nonlinear approach to endmember identification, etc...

@@

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}
}

>>>>

1. Water quality classification

Inland water bodies pose a unique challenge for classification by remote sensing imagery due to their complex spectral characteristics, which stem from a combination of chemical, biological, and environmental factors.

The complex spectral characteristics of inland water bodies pose a unique challenge to their classification by remote sensing methods due to the combination of chemical, biological, and environmental sources. Remote sensing applications for water quality include monitoring optically-active parameters like turbidity, temperature, chlorophyll-a, blue-green algae, etc. \cite{ritchie2003remote}. Spectral indices together with machine learning and simulation remain key approaches \cite{sagan2020monitoring}. Remote sensing has been used for analysis of oil spills such as the 2010 Deepwater Horizon spill \cite{leifer2012state,kokaly2013spectroscopic}. Another application is for the identification and classification of harmful algal blooms (HABs) \cite{thenkabail2018hyperspectral,ghatkar2019classification}. The application of remote sensing to water quality monitoring is largely focused on the estimation of optically-active constituents such as chlorophyll-a, blue-green algae, temperature, and turbidity \cite{ritchie2003remote}.

2. Limitations of traditional remote sensing: Limited spectral resolution and need for expensive in-situ reference data

A key limitation is the need for sufficient quantities of in-situ reference data. Aurin et al. took a similar route combining 30 years of remote sensing imagery with in-situ data from over 500 field compaigns for CDOM, organic carbon, etc... \cite{aurin2018remote}. Ross et al. created a comprehensive dataset with over 600,000 water quality records matching optically active water quality parameters with associated satellite imagery from Landsat 5,7,8. To achieve this quantity of data, they needed records spanning 1984 to 2019 \cite{ross2019aquasat}.

3. Data Driven methods for classification

When high resolution ground truth data are not available, one approach is to classify water quality into a set of categories ranging from bad to good \cite{koponen2002lake}. Many data-driven, unsupervised classification techniques have been employed for this task for remotely sensed multi-spectral and hyperspectral imagery including various matrix factorization methods, k-nearest neighbors, fuzzy c-means, density estimation, etc... \cite{zhang2019hyperspectral}. Additionally, ICA, LDA, and similar techniques can be used to perform nonlinear dimensionality reduction (as opposed to PCA) for visualization or as a preprocessing step for supervised regression and classification tasks \cite{liang2010hyperspectral, zhangSVM2014}.

4. Hyperspectral imaging addresses the first limitation by providing improved spectral resolution

Hyperspectral data can help distinguish oils by type, e.g. crude, diesel, gasoline, and palm oil \cite{yang2020characterization}. Other applications of hyperspectral imaging include food quality, food safety, medical diagnosis, precision agriculture, and forensic document examination \cite{khan2018modern}

5.

1. Hyperspectral imaging is a powerful tool. Specifically for water quality, the increased spectral resolution enables the identification and discernment of a variety of contaminants from harmful algal blooms (HABs) to oil spills which pose a significant challenge to traditional multi-spectral remote sensing systems...

2. Traditionally, the applications of such imagers to water quality monitoring relies on ground-truth observations using reference instruments. For example... chorophyll-a, blue-green algae, crude oil, etc. However reference instruments are expensive to deploy and challenging to maintain often requiring complicated calibration

3. Unsupervised classification of hyperspectral imagery. Discuss ML approaches including Self Organizing Map (SOM).

4. A related involves the data-driven identification of unique spectral constituents within a scene known as endmember extraction. A variety of statistical and machine learning based approaches have been developed to extract unique spectral signatures corresponding to different sources. Once these libraries of spectral signatures are obtained, they can be combined to model the spectrum of each pixel in a HSI. This approach is known as spectral unmixing and has been used to great success for a variety of applications including

4. While HSI offer superior spectral resolution, the limited spatial and temporal resolution provided by satellite-based platforms remains a challenge for the analysis of inland waters with small-scale spatial variability. Additionally, the infrequency of satellite overpasses makes it challenging to identify sources of contamination or spills in real time. With the continual development of hyperspectral imaging technology, the incorporation of hyperspectral imagers as the payload of autonomous aerial vehicles (UAVs) such as quadcopters, hexacopters, and other drones offers unparalleled capabilities with nearly centimeter accuracy.

5. Robot team

6. In this paper, we demonstrate how Generative Topographic Mapping (GTM), a Bayesian alternative to the Self Organizing Map, can be utilized to rapidly identify relevant spectral signatures within a scene. Using data collected at site in Montague, North Texas, we demonstrate the ability of this paradigm with two case studies. First, we demonstrate how the GTM can be used to provide a high-resolution unsupervised classification of a water body enabling the rapid identification of relevant regions which can then be used to direct further investigation by reference-sensor equipped boats. Second, we demonstrate the ability to perform endmember extraction and abundance mapping using imagery from a test-release of rhodamine dye. Together with the normalized spectral similarity score, these spectral signatures are identified and their associated area in the scene is determined.

@@

\noindent 
\begin{tabular}{@{}ll}
HAB & Harmful Algal Bloom \\ 
HSI & Hyperspectral Image \\ 
SOM & Self Organizing Map \\ 
GTM & Generative Topographic Mapping 
\end{tabular}
}
